---
title: "Extract series from an ensemble of RCMs for a station"
author: "Michael Matiu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extract series from an ensemble of RCMs for a station}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The goal: Extract temperature (tas) and precipitation (pr) of a single grid cell for the SMHI-RCA4 regional climate model: it will look at all available GCMs and we subset to historical, rcp45 and rcp85.

The single grid cell will be the one which includes the station of Bolzano (Italy), which lies at 11.31 deg E and 46.5 deg N.

First get inventory of all files.

```{r}
library(eurocordexr)

path_eurocordex <- "/mnt/CEPH_BASEDATA/METEO/SCENARIO"
dat_inventory <- get_inventory(path_eurocordex)
dat_inventory_files <- get_inventory(path_eurocordex, add_files = T)

# the inventory without files prints more nicely
dat_inventory
```

Then we subset using data.table notation:

```{r}
dat_inventory_sub <- dat_inventory[variable %in% c("tas", "pr") & 
                                     timefreq == "day" & 
                                     institute_rcm == "SMHI-RCA4" &
                                     experiment %in% c("historical", "rcp45", "rcp85")]
# the inventory with files is needed later for data extraction
dat_inventory_files_sub <- dat_inventory_files[variable %in% c("tas", "pr") & 
                                     timefreq == "day" & 
                                     institute_rcm == "SMHI-RCA4" &
                                     experiment %in% c("historical", "rcp45", "rcp85")]
```


Next compare the available model setups:

```{r}
dat_compare <- compare_variables_in_inventory(dat_inventory_sub, c("tas", "pr"))
print(dat_compare)
```

Looks like it is a complete dataset. Now we are going to extract all the data for the grid cell in which the station lies.

We go parallel, because it saves a lot of time (for the example here, we use 7 instead of 1 core).


```{r}

# get parallel
library(doParallel)
registerDoParallel(parallel::detectCores() - 1)

```

The extraction of the data is enclosed in the following loop. It will create one file (.rds) per row in the inventory data. In this case it is one file per variable (tas, pr), model (GCM - RCM combination), and experiment (historical, rcp..). The data will be saved in a directory `data-rcm/[station-name]/[variable]`, which will be created if it does not exist. We will also, for illustration, measure the time it takes to extract all data.



```{r}

# time the following loop
start_time <- Sys.time()

# extract data and save files
library(foreach)
foreach(
  i_stn = "Bolzano",
  i_lon = 11.31,
  i_lat = 46.5
) %do% {
  
  foreach(i_nc = 1:nrow(dat_inventory_files_sub), .inorder = F) %dopar% {
    
    # create filename
    dat_inventory_files_sub[i_nc, paste0(# variable, "_",
                         gcm, "_",
                         institute_rcm, "_",
                         experiment, "_",
                         ensemble, "_",
                         downscale_realisation, ".rds")] %>% 
      # pre-append path
      file.path("data-rcm", 
                i_stn, 
                dat_inventory_files_sub[i_nc, variable], 
                .) -> file_to_save
    
    # create directory and skip if file already exists
    if(!dir.exists(dirname(file_to_save))) dir.create(dirname(file_to_save), recursive = T)
    if(file.exists(file_to_save)) return(NULL)

    # subloop to extract all data for a specific inventory row
    dat_nc <- foreach(
      fn = dat_inventory_files_sub[i_nc, unlist(list_files)],
      .final = rbindlist
    ) %do% {
      
      dat <- rotpole_nc_point_to_dt(filename = fn,
                                    variable = dat_inventory_files_sub[i_nc, variable],
                                    point_lon = i_lon,
                                    point_lat = i_lat,
                                    interpolate_to_standard_calendar = T,
                                    verbose = F)
      
      dat
      
      
    }
    
    saveRDS(dat_nc,
            file = file_to_save)
    
    return(NULL)
  }
  
  return(NULL)
}

# how much time has passed?
end_time <- Sys.time()
print(end_time - start_time)

```

It took only approx. 8 minutes to extract all the data.

The above function can also be adopted to loop over multiple stations, just giving in the first `foreach()` call vectors of `i_stn`, `i_lon`, and `i_lat` (which can also be columns from a data.frame/data.table). 
