---
title: "Merge extracted station data"
author: "Michael Matiu"
date: "5/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this documentation, some example code will be given to merge the data, which has been extracted as in the separate documentation.


We assume the folder structure below and the associated combination fucntion: merge for variables (pr, tas), and rbindlist for stations and single models.

```{r}
library(foreach)
library(data.table)
library(magrittr)
library(ggplot2)

# main dir --- stn       --- var   --- *.rds
#              rbindlist --- merge --- rbindlist

dat_full <- foreach(
  i_dir_stn = list.dirs("data-rcm", recursive = F),
  .final = rbindlist
) %do% {
  
  foreach(
    i_dir_stn_var = list.dirs(i_dir_stn, recursive = F),
    .final = function(l) Reduce(merge, l)
  ) %do% {
    
    foreach(
      fn = list.files(i_dir_stn_var, full.names = T, recursive = F),
      .final = rbindlist
    ) %do% {
      
      dat <- readRDS(fn)
      fn_split <- strsplit(basename(fn), "_")[[1]]
      dir_split <- strsplit(dirname(fn), "/")[[1]]
      
      dat[, ":="(stn = dir_split[2],
                 gcm = fn_split[1],
                 rcm = fn_split[2],
                 experiment = fn_split[3],
                 ensemble = fn_split[4])]
      
      dat
      
    }
    
  }
  
}
```

Next, we convert the units to mm for precipitation and degree Celsius for temperature.
```{r}
# C and mm
dat_full[, ":="(pr = pr * 24*60*60, tas = tas - 273.15)]
```


For fun, now we just calculate the seasonal changes in temperature and precipitation for the different rcp scenarios.

```{r, fig.width=12, fig.height=7}
# add seasons
seasons <- c("DJF", "DJF", rep(c("MAM", "JJA", "SON"), each = 3), "DJF")
dat_full[, season := seasons[month(date)]]
dat_full[, season := factor(season, levels = c("DJF", "MAM", "JJA", "SON"))]
dat_full[, seasonyear := year(date)]
dat_full[month(date) == 12, seasonyear := seasonyear + 1L]

# seasonal  means/sum
dat_mean_season <- dat_full[seasonyear >= 1971 & seasonyear <= 2000 |
                              seasonyear >= 2070 & seasonyear <= 2099,
                            .(nn = .N, sum_pr = sum(pr), mean_tas = mean(tas)),
                            .(season, seasonyear, stn, gcm, rcm, experiment, ensemble)]

# climatological means
dat_mean_clim <- dat_mean_season[,
                                 .(mean_sum_pr = mean(sum_pr),
                                   mean_mean_tas = mean(mean_tas)),
                                 .(season, stn, gcm, rcm, experiment, ensemble)]

# calculate deltas
dat_mean_clim %>% 
  dcast(season + stn + gcm + rcm + ensemble ~ experiment, value.var = "mean_mean_tas") %>% 
  .[, .(season, stn, gcm, rcm, ensemble,
        # rcp26 = rcp26 - historical,
        rcp45 = rcp45 - historical,
        rcp85 = rcp85 - historical)] %>% 
  melt(measure.vars = patterns("rcp"), 
       variable.name = "experiment",
       value.name = "delta_tas") -> dat_delta_tas

dat_mean_clim %>% 
  dcast(season + stn + gcm + rcm + ensemble ~ experiment, value.var = "mean_sum_pr") %>% 
  .[, .(season, stn, gcm, rcm, ensemble,
        # rcp26 = rcp26 - historical,
        rcp45 = rcp45 - historical,
        rcp85 = rcp85 - historical)] %>% 
  melt(measure.vars = patterns("rcp"), 
       variable.name = "experiment",
       value.name = "delta_pr") -> dat_delta_pr


# plot
dat_plot <- merge(dat_delta_tas, dat_delta_pr)

dat_plot %>% 
  ggplot(aes(delta_tas, delta_pr, colour = rcm, shape = gcm))+
  geom_point(size = 3)+
  facet_grid(experiment~season)+
  theme_bw()+
  ggtitle("Climate models (RCM+GCM combination) spread in projected changes 2070-2099 (rel. to 1971-2000)")+
  xlab("Delta Temperature [deg C]")+
  ylab("Delta precipitation [mm]")
  
```
