---
title: "Reading a rotated-pole grid into R"
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eurocordexr)
library(data.table)

# for plotting
library(ggplot2)
library(scico)
```


This will show how to read in a grid in long data.table format using the native rotated pole grid.

Note that climate model data can be very heavy, so I like to separate actual data from coordinates and do much of the pre-processing with CDO. For the example here, it's not that relevant, since file sizes are small - but it can be useful if you have to do a lot of merging or aggregating.

Naturally, the long format used in eurocordexr is not the best choice for array data, but it is more user-friendly when merging and aggregating using data.table or dplyr syntax. And for plotting with ggplot. 


The example data (created with CDO) within the package is seasonal means for 1971-2000 and 2071-2100 under RCP8.5 for REMO2015 driven by NorESM, cropped to the European Alps.

```{r}
file_nc_past <- system.file("extdata", "alps1.nc", package = "eurocordexr")
file_nc_future <- system.file("extdata", "alps2.nc", package = "eurocordexr")
```


## Staying in the native rotated-pole grid (with eurocordexr)

First we get the coordinates (just one timestep):

```{r}
dat_aux <- nc_grid_to_dt(file_nc_past, add_xy = T, date_range = c("2000-01-15", "2000-01-15"))
# keep only cell index (for merging) and coordinates (which have some rounding issues)
dat_aux <- dat_aux[, .(icell, rlon = round(rlon, 3), rlat  = round(rlat, 3))]
```

Then the actual data.

```{r}
dat_past <- nc_grid_to_dt(file_nc_past)
dat_future <- nc_grid_to_dt(file_nc_future)
```

For nicer plotting, create a factor of season labels, with helper function from https://github.com/mitmat/mitmatmisc/.


```{r}
dat_past[, season := mitmatmisc::season_fct(month(date))]
dat_future[, season := mitmatmisc::season_fct(month(date))]
```



And finally plot the past climatology.

```{r}
dat_past |> 
  merge(dat_aux) |> 
  ggplot(aes(rlon, rlat, fill = tas))+
  geom_raster()+
  facet_wrap(~season)+
  theme_minimal()+
  scale_fill_scico(direction = -1)
```

Or the future change signal.

```{r}
dat_past[, .(icell, season, tas_past = tas)] |> 
  merge(dat_future[, .(icell, season, tas_future = tas)]) |> 
  merge(dat_aux, by = "icell") |> 
  
  ggplot(aes(rlon, rlat, fill = tas_future - tas_past))+
  geom_raster()+
  facet_wrap(~season)+
  theme_minimal()+
  scale_fill_scico(palette = "vik", midpoint = 0)
```


## Using a curvilinear grid (stars package)

The stars package has some support for the rotated-pole grid with its capability to read curvilinear grids. This won't give you the best performance, but can work if you do much of the preprocessing outside R.


```{r}
library(stars)
rs <- read_stars(file_nc_past)
#  take 1 timestep
rs1 <- rs[,,,1]

plot(rs1)
```

Or with ggplot:

```{r}
ggplot()+
  geom_stars(data = rs1)+
  theme_minimal()+
  scale_fill_scico(direction = -1)
```


## Using a regular lonlat grid (terra package)

Finally, the third way is to first convert the rotated-pole grid into a regular lonlat grid. See the [other article](rotpole-cdo.html) on how to do it with CDO. 

```{r}
library(terra)
file_nc_regular <- system.file("extdata", "alps1_regular.nc", package = "eurocordexr")
rs2 <- rast(file_nc_regular)
plot(rs2)
```







